---
title: "Chapter 4"
output: html_notebook
---
# Programing basics

## 4.1 Conditional expressions

Conditional expressions은 flow control에 주로 사용된다. 가장 일반적인 conditional expressions은 if-else이다.
```{r}
#a가 0이 아니면 1/a를 print하고 그렇지 않다면 "No reciprocal for 0"을 print하라.
a <- 0

if(a!=0){
  print(1/a)
} else{
  print("No reciprocal for 0.")
}
```

Let’s look at one more example using the US murders data frame:
```{r}
library(dslabs)
data(murders)
murder_rate <- murders$total / murders$population*100000
```

```{r}
#murder_rate가 0.5보다 작은 state가 있다면 minimum값을 가지는 state를 print하고, 없다면 "No state has murder rate that low"를 print 하라.

ind <- which.min(murder_rate)

if(murder_rate[ind] < 0.5){
  print(murders$state[ind]) 
} else{
  print("No state has murder rate that low")
}
```

murder rates를 0.25보다 작게 설정하면 결과는 달라진다.
```{r}
if(murder_rate[ind] < 0.25){
  print(murders$state[ind]) 
} else{
  print("No state has a murder rate that low.")
}
```

이와 관련하여 `ifelse`를 사용할 수도 있다. This function takes three arguments: a logical and two possible answers. If the logical is `TRUE`, the value in the **second** argument is returned and if `FALSE`, the value in the **third** argument is returned. 
```{r}
a <- 0
ifelse(a>0, 1/a, NA)
```

The function is particularly useful because it works on vectors.
```{r}
a <- c(0, 1, 2, -4, -5)
result <- ifelse(a>0, 1/a, NA)
result
```

Here is an example of how this function can be readily used to replace all the missing values in a vector with zeros:
```{r}
data(na_example)
no_nas <- ifelse(is.na(na_example), 0, na_example) 
sum(is.na(no_nas))
```

또다른 function으로는 `any`와 `all`이 있다. `any` takes a vector of logicals and returns `TRUE` if any of the entries is `TRUE`. `all` takes a vector of logicals and returns `TRUE` if all of the entries are `TRUE`. 
```{r}
z <- c(TRUE, TRUE, FALSE)
any(z)
all(z)
```

## 4.2 Defining functions

특정 function을 반복적으로 사용할 때, 이를 직접 지정하여 사용할 수 있다. 예를들어 vector `x`의 average를 알고자 할 때, `sum`과 `length`를 이용하여 sum(x)/length(x) 값을 구한다. 결과적으로 average를 구하는 기능은 보편적으로 많이 쓰기 때문에 R에서 `mean`이라는 function으로 따로 지정하여 두었지만 지정된 function이 없을 경우에는 따로 정의내려 사용하는 것이 편하다.
```{r}
avg <- function(x){
  s <- sum(x)
  n <- length(x)
  s/n
}
```

```{r}
#average와 mean은 결과적으로 같은 의미
x <- 1:100
identical(mean(x), avg(x))
```

Notice that variables defined inside a function are not saved in the workspace. So while we use `s` and `n` when we call `avg`, the values are created and changed only during the call.
```{r}
s <- 3
avg(1:10)
s
```
이미 `s`를 3으로 지정했는데 어떻게 `avg` function은 그대로 유지될 수 있는지?

The general form of a function definition looks like this:
```{r}
#General form of a function definition

my_function <- function(VARIABLE_NAME){
  perform operations on VARIABLE_NAME and calculate VALUE
  VALUE
}
```

The functions you define can have multiple arguments as well as default values. For example, we can define a function that computes either the arithmetic or geometric average.
```{r}
avg <- function(x, arithmetic = TRUE){
  n <- length(x)
  ifelse(arithmetic, sum(x)/n, prod(x)^(1/n))
}
```

## 4.3 Namespaces

We will likely need to load several add-on packages for some of your analysis. It is likely that two packages use the same name for two different functions. And often these functions do completely different things. 
```{r}
The following objects are masked from ‘package:stats’:

    filter, lag

The following objects are masked from ‘package:base’:

    intersect, setdiff, setequal, union
```

R will follow a certain order when searching for a function in these *namespaces*.
```{r}
search()
```

만약 **stats** `filter`을 사용하고 싶다면,
```{r}
stats::filter
```

만약 **dplyr** `filter`을 사용하고 싶다면.
```{r}
dplyr::filter
```

만약 package 전체를 로딩하지 않고 function을 사용하고 싶다면 `::`을 이용할 수 있다.