---
title: "Chapter 3"
output: html_notebook
---

## 3.2. The very basics

### 3.2.1 Objects


수식을 해결할 때 e.g. "ax^2+bx+c=0", 

근의 공식[{-b-(b^2 -4ac)^1/2}/2a and {-b+(b^2 -4ac)^1/2}/2a]을 이용,  a, b, c를 지정해준다.

To solve x^2+x-1=0,  a, b, c에 해당하는 숫자를 지정해준다.
```{r}
a <-1
b <- 1
c <- -1
```

To show the value stored in 'a' is using 'print':
```{r}
print(a)
```

### 3.2.2 The workspace

When I want to see all the variables saved in my workspace, use `ls`
```{r}
ls()
```
만약 아직 지정하지 않은 value를 입력한다면 error code가 뜬다.

위의 value는 이미 저장되었으므로 방정식의 해를 얻기 위해서 근의 공식 사용:
```{r}
(-b + sqrt(b^2 - 4*a*c) ) / ( 2*a )
(-b - sqrt(b^2 - 4*a*c) ) / ( 2*a )
```

### 3.2.3 Functions

어떠한 기능을 알고 싶을 때 `help("")` 이용, e.g. log:
```{r}
help("log")
```

코드에 어떤 항목을 입력해야 하는지 알고 싶을 때 `args()` 이용, e.g. log:
```{r}
args(log)
```
for example,
```{r}
log(8, base=2)
log(8, 2)
```

순서를 바꿔쓰고 싶다면 문자를 지정해주면 된다.
```{r}
log(base=2, x=8)
```

### 3.2.4 Other prebuilt objects

data를 보고 싶을 때 `data` 입력
```{r}
data()
```

### 3.2.5 Variable names

For the quadratic equations, we could use something like this:
```{r}
solution_1 <- (-b + sqrt(b^2 - 4*a*c)) / (2*a)
solution_2 <- (-b - sqrt(b^2 - 4*a*c)) / (2*a)
```

### 3.2.6 Saving your workspace

### 3.2.7 Motivating scripts

만약 다른 방정식 e.g. 3x^2+2x-1 을 풀고 싶다면 value에 해당하는 값을 다시 지정해주면 된다.
```{r}
a <- 3
b <- 2
c <- -1
(-b + sqrt(b^2 - 4*a*c)) / (2*a)
(-b - sqrt(b^2 - 4*a*c)) / (2*a)
```

### 3.2.8 Commenting your code

code block 부분에 #을 입력하면 code로 인식하지 않는다.
```{r}
## Code to compute solution to quadratic equation of the form ax^2 + bx + c
## define the variables
a <- 3 
b <- 2
c <- -1

## now compute the solution
(-b + sqrt(b^2 - 4*a*c)) / (2*a)
(-b - sqrt(b^2 - 4*a*c)) / (2*a)
```

## 3.3 Exercises

1. What is the sum of the first 100 positive integers? The formula for the sum of integers 1 through n is n(n+1)/2. Define n=100 and then use R to compute the sum of 1 through 100 using the formula. What is the sum?
```{r}
#따로 1부터 n까지 나열하는 방법을 모른다면, n에 원하는 숫자 지정한 후, 이용하고자 하는 수식을 그대로 입력한다.
n<-100
n*(n+1)/2
```
2. Now use the same formula to compute the sum of the integers from 1 through 1,000
```{r}
n<-1000
n*(n+1)/2
```
3. Look at the result of typing the following code into R:
```{r}
n<-1000
x<-seq(1,n)
sum(x)
#seq(1,n)은 1부터 n까지 나열하는 명령어, sum(x)는 x에 지정된 것을 모두 더하는 것
```
4. In math and programming, we say that we evaluate a function when we replace the argument with a given number. So if we type `sqrt(4)`, we evaluate the `sqrt` function. In R, you can evaluate a function inside another function. The evaluations happen from the inside out. Use one line of code to compute the log, in base 10, of the square root of 100.


5. Which of the following will always return the numeric value stored in 'x'? You can try out examples and use the help system if you want.
```{r}
log(10^x)
log10(x^10)
log(exp(x))
exp(log(x, base=2))

#exp는 자연함수를 지칭하고 R에서의 log는 일반적으로 밑이 e로 설정되어 있다.
```

## 3.4 Data types

Variables in R can be of different types. For example, we need to distinguish numbers from character strings and tables from simple lists of numbers. The function `class` helps us determine what type of object we have:

```{r}
a <- 2
class(a)
```

## 3.5 Data frames

Storing a dataset in R is in a data frame. Data analysis challenges start with data stored in a data frame.
```{r}
install.packages("dslabs")
library(dslabs)
data(murders)
#Data frame을 불러오기 위해 `class` 이용
class(murders)
```

### 3.5.1 Examining an object

`str`은 object의 전반적인 structure을 알 수 있다.
```{r}
str(murders)
```

`head`는 가지고 있는 data의 처음 6줄 data를 보여준다.
```{r}
head(murders)
```

### 3.5.2 The accessor: `$`

Access the different variables represented by columns included in this data frame.
```{r}
murders$population
```

Data frame을 구성하는 항목을 알고 싶다면 `names`을 이용
```{r}
names(murders)
```

### 3.5.3 Vectors: numerics, characters, and logical

`length`는 몇 개의 entries가 vector에 있는지 보여준다.
```{r}
pop <- murders$population
length(pop)
```
This particular vector is numeric since population sizes are numbers:
```{r}
class(pop)
```
To store character strings, vectors can also be of class character.
```{r}
class(murders$state)
```

Logical vectors `==`: true 혹은 false로 결과값을 보여줌
```{r}
z <- 3==2
z
```

### 3.5.4 Factors

'murders' dataset에서 region도 마찬가지로 character vector일 것이라 생각하지만 그렇지 않다.
```{r}
class(murders$region)
```

Factors are useful for storing categorical data. We can see that there are only 4 regions by using the `levels` function:
```{r}
levels(murders$region)
```

The function `reorder` lets us change the order of the levels of a factor variable based on a summary computed on a numeric vector.
```{r}
region <- murders$region
value <- murders$total
region <- reorder(region, value, FUN = sum)
levels(region)
```

### 3.5.5 Lists

Data frames are a special case of lists. We can store any combination of different types.

### 3.5.6 Matrices

We can define a matrix using the `matrix` function. 
```{r}
mat <- matrix(1:12, 4, 3)
mat
```

We can access specific entries in a matrix.
```{r}
#e.g. 2행 3열 entry를 알고 싶다면,
mat[2,3]
```

```{r}
#e.g. 2행의 모든 entries를 알고 싶다면,
mat[2,]
```

```{r}
#e.g. 3열의 모든 entries를 알고 싶다면,
mat[,3]
```

We can access more than one column or more than one row.
```{r}
#e.g. 2열과 3열의 entries를 알고 싶다면,
mat[,2:3]
```

```{r}
#e.g. 1행과 2행 그리고 2열과 3열의 entries를 알고 싶다면,
mat[1:2,2:3]
```

`as.data.frame` function을 이용하면 matrices를 data frames으로 변환할 수 있다.
```{r}
as.data.frame(mat)
```

Data frame에서도 matrices와 마찬가지로 []을 이용하여 특정 행과 열의 값을 알 수 있다.
```{r}
data("murders")
murders[25, 1]
murders[2:3, ]
```

## 3.6 Excercises

1. Load the US murders dataset.
```{r}
library(dslabs)
data(murders)
```

Use the function `str` to examine the structure of the 'murders' object. We can see that this object is a data frame with 51 rows and five columns. Which of the following best describes the variables represented in this data frame?
```{r}
str(murders)
```

C. The state name, the abbreviation of the state name, the state’s region, and the state’s population and total number of murders for 2010.

2. What are the column names used by the data frame for these five variables?
```{r}
names(murders)
```

3. Use the accessor `$` to extract the state abbreviations and assign them to the object `a`. What is the class of this object?
```{r}
murders$abb
a <- murders$abb
class(a)
```

4. Now use the square brackets to extract the state abbreviations and assign them to the object `b`. Use the `identical` function to determine if a and `b` are the same
```{r}
data("murders")
b <- murders[,2]
a==b
```

5. We saw that the `region` column stores a factor. You can corroborate this by typing:
```{r}
class(murders$region)
levels(region)
length(region)
```

6. The function `table` takes a vector and returns the frequency of each element. You can quickly see how many states are in each region by applying this function. Use this function in one line of code to create a table of states per region.
```{r}
table(region)
```
