---
title: "**Chapter 5. The tidyverse**"
output: html_notebook
---

## **5.7 Summarizing data**

### **5.7.1 `summarize`**

The `summarize` function in **dplyr** provides a way to compute summary statistics with intuitive and readable code.
```{r}
library(dplyr)
library(dslabs)
data(heights)
```

```{r}
s <- heights %>% 
  filter(sex == "Female") %>%
  summarize(average = mean(height), standard_deviation = sd(height))
s
```
`s`는 결과적으로 data frame이기 때문에 `$`를 이용해서 data에 접근할 수도 있다.
```{r}
s$average
s$standard_deviation
```

`mean`값 외에도 `median`, `minimum`, `maximum`을 살펴볼 수도 있다.
```{r}
heights %>% 
  filter(sex == "Female") %>%
  summarize(median = median(height), minimum = min(height), maximum = max(height))
```

We can obtain these three values with just one line using the `quantile` function: for example, `quantile(x, c(0,0.5,1))` returns the min (0th percentile), median (50th percentile), and max (100th percentile) of the vector `x`.
```{r}
heights %>% 
  filter(sex == "Female") %>%
  summarize(range = quantile(height, c(0, 0.5, 1)))
```
하지만 `summarize` function에서는 single value만 return되기 때문에 error code가 발생한다. 

*Another example*
```{r}
data(murders)
murders <- murders %>% mutate(rate = total/population*100000)
summarize(murders, mean(rate))
```
The US murder rate is not the average of the state murder rates. This is because in the computation above the small states are given the same weight as the large ones. The US murder rate is the total number of murders in the US divided by the total US population.
```{r}
us_murder_rate <- murders %>%
  summarize(rate=sum(total)/sum(population)*100000)
us_murder_rate
```

### **5.7.2 `pull`**
```{r}
class(us_murder_rate)
```
##### `summarize` always returns a data frame.

When a data object is piped that object and its columns can be accessed using the `pull` function.
```{r}
us_murder_rate %>% pull(rate)
```
This returns the value in the rate column of `us_murder_rate` making it equivalent to `us_murder_rate$rate`.

To get a number from the original data table with one line of code we can type:
```{r}
us_murder_rate <- murders %>% 
  summarize(rate = sum(total) / sum(population) * 100000) %>%
  pull(rate)
us_murder_rate
```
```{r}
class(us_murder_rate)
```

### **5.7.3 Group then summarize with `group_by`**

A common operation in data exploration is to first split data into groups and then compute summaries for each group.
```{r}
heights %>% group_by(sex)
```
Although not immediately obvious from its appearance, this is now a special data frame called a grouped *data frame* and **dplyr** functions, in particular `summarize`, will behave differently when acting on this object.
```{r}
heights %>% 
  group_by(sex) %>%
  summarize(average = mean(height), standard_deviation = sd(height))
```
*Another example*
```{r}
murders %>% 
  group_by(region) %>%
  summarize(median_rate = median(rate))
```

## **5.8 Sorting data frames**

We know about the `order` and `sort` function, but for ordering entire tables, the dplyr function `arrange` is useful. 
```{r}
murders %>%
  arrange(population) %>%
  head()
```
With `arrange` we get to decide which column to sort by. To see the states by population, from smallest to largest, we arrange by `rate` instead:
```{r}
murders %>% 
  arrange(rate) %>% 
  head()
```
In dplyr, the function `desc` transforms a vector so that it is in descending order.
```{r}
murders %>% 
  arrange(desc(rate)) %>% 
  head()
```

### **5.8.1 Nested sorting**

Here we order by `region`, then within region we order by murder rate:
```{r}
murders %>% 
  arrange(region, rate) %>% 
  head()
```

### **5.8.2 The top *n* **

If we want to see a larger proportion, we can use the `top_n` function. This function takes a data frame as it’s first argument, the number of rows to show in the second, and the variable to filter by in the third.
```{r}
murders %>% top_n(10, rate)
```
여기서 기억해야 할 사실은 row는 `rate`로 *filter*되었을 뿐 *sort*되지는 않았다는 것이다. 추가로 `rate`로 sorting하고 싶으면 `arrange`를 사용하면 된다.

## **5.9 Exercises**

For these exercises, we will be using the data from the survey collected by the United States National Center for Health Statistics (NCHS). This center has conducted a series of health and nutrition surveys since the 1960’s. Starting in 1999, about 5,000 individuals of all ages have been interviewed every year and they complete the health examination component of the survey. Part of the data is made available via the **NHANES** package. Once you install the **NHANES** package, you can load the data like this:
```{r}
install.packages("NHANES")
library(NHANES)
data(NHANES)
```
The **NHANES** data has many missing values. Remember that the main summarization function in R will return `NA` if any of the entries of the input vector is an `NA`. Here is an example:
```{r}
library(dslabs)
data(na_example)
mean(na_example)
sd(na_example)
```
To ignore the `NA`s we can use the `na.rm` argument:
```{r}
mean(na_example, na.rm=TRUE)
sd(na_example, na.rm=TRUE)
```
Let’s now explore the NHANES data.

**1. We will provide some basic facts about blood pressure. First let’s select a group to set the standard. We will use 20-29 year old females. `AgeDecade` is a categorical variable with these ages. Note that the category is coded like " 20-29“, with a space in front! What is the average and standard deviation of systolic blood pressure as saved in the `BPSysAve` variable? Save it to a variable called `ref`.**

*Hint: Use `filter` and `summarize` and use the `na.rm = TRUE` argument when computing the average and standard deviation. You can also filter the NA values using `filter`.*
```{r}
ref <- NHANES %>%
  filter(Gender == "female") %>%
  filter(AgeDecade == " 20-29") %>%
  summarize(average=mean(BPSysAve, na.rm = TRUE), standard_deviation= sd(BPSysAve, na.rm = TRUE))
ref
```

**2. Using a pipe, assign the average to a numeric variable `ref_avg`. Hint: Use the code similar to above and then `pull`.**
```{r}
ref_avg <- NHANES %>%
  filter(Gender == "female") %>%
  filter(AgeDecade == " 20-29") %>%
  mean(BPSysAve, na.rm = TRUE)
```

**3. Now report the min and max values for the same group.**
```{r}
NHANES %>%
  filter(Gender == "female") %>%
  filter(AgeDecade == " 20-29") %>%
  summarize(minimum=min(BPSysAve, na.rm = TRUE), maximum=max(BPSysAve, na.rm = TRUE))
```

**4. Compute the average and standard deviation for females, but for each age group separately rather than a selected decade as in question 1. Note that the age groups are defined by `AgeDecade`. Hint: rather than filtering by age and gender, filter by `Gender` and then use `group_by`.**
```{r}
NHANES %>%
  filter(Gender == "female") %>%
  group_by(AgeDecade) %>%
  summarise(average=mean(BPSysAve, na.rm = TRUE), standard_deviation=sd(BPSysAve, na.rm =TRUE))
```

**5. Repeat exercise 4 for males.**
```{r}
NHANES %>%
  filter(Gender == "male") %>%
  group_by(AgeDecade) %>%
  summarise(average=mean(BPSysAve, na.rm = TRUE), standard_deviation=sd(BPSysAve, na.rm =TRUE))
```

**6. We can actually combine both summaries for exercises 4 and 5 into one line of code. This is because `group_by` permits us to group by more than one variable. Obtain one big summary table using `group_by(AgeDecade, Gender)`.**
```{r}
NHANES %>%
  group_by(AgeDecade, Gender) %>%
  summarise(average=mean(BPSysAve, na.rm = TRUE), standard_deviation=sd(BPSysAve, na.rm =TRUE))
```

**7. For males between the ages of 40-49, compare systolic blood pressure across race as reported in the `Race1` variable. Order the resulting table from lowest to highest average systolic blood pressure.**
```{r}
NHANES %>%
  filter(Gender == "male") %>%
  filter(AgeDecade == " 40-49") %>%
  group_by(Race1) %>%
  summarize(average=mean(BPSysAve, na.rm = TRUE)) %>%
  arrange(average)
```

## **5.10 Tibbles**

Tidy data must be stored in data frames. We introduced the data frame in Section 3.5 and have been using the `murders` data frame throughout the book:
```{r}
data(murders)
class(murders)
```

In Section 5.7.3 we introduced the `group_by` function, which permits stratifying data before computing summary statistics.
```{r}
murders %>% group_by(region) %>% head()
```
```{r}
murders %>% group_by(region) %>% class()
```
The `tbl`, pronounced tibble, is a special kind of data frame. The functions `group_by` and `summarize` always return this type of data frame. The group_by function returns a special kind of `tbl`, the `grouped_df`.

### **5.10.1 Tibbles display better**

The print method for tibbles is more readable than that of a data frame. To see this, compare the outputs of typing `murders` and the output of murders if we convert it to a tibble. We can do this using `as_tibble(murders)`.
```{r}
murders
as_tibble(murders)
```

### **5.10.2 Subsets of tibbles are tibbles**

If you subset the columns of a data frame, you may get back an object that is not a data frame, such as a vector or scalar.
```{r}
class(murders[,4])
```
is not a data frame. With tibbles this does not happen:
```{r}
class(as_tibble(murders)[,4])
```
With tibbles, if you want to access the vector that defines a column, and not get back a data frame, you need to use the accessor `$`:
```{r}
class(as_tibble(murders)$population)
```
If we try this with a tibble we get an informative warning:
```{r}
as_tibble(murders)$Population
```

### **5.10.3 Tibbles can have complex entries**

While data frame columns need to be vectors of numbers, strings or logical values, tibbles can have more complex objects, such as lists or functions. Also, we can create tibbles with functions:
```{r}
tibble(id=c(1, 2, 3), func=c(mean, median, sd))
```

### **5.10.4 Tibbles can be grouped**

The function `group_by` returns a special kind of tibble: a grouped tibble. This class stores information that lets you know which rows are in which groups.

### **5.10.5 Create a tibble using `date_frame` instead of `data.frame`**

To create a data frame in the tibble format, you can do this by using the `data_frame` function.
```{r}
grades <- data_frame(names=c("John", "Juan", "Jean", "Yao"),
                     exam_1=c(95, 80, 90, 95),
                     exam_2=c(90, 85, 85, 90))
grades
```
`data.frame` can be used to create a regular data frame rather than a tibble. One other important difference is that by default `data.frame` coerces characters into factors without providing a warning or message:
```{r}
grades <- data.frame(names = c("John", "Juan", "Jean", "Yao"), 
                     exam_1 = c(95, 80, 90, 85), 
                     exam_2 = c(90, 85, 85, 90))
class(grades$names)
```
To avoid this, we use the rather cumbersome argument `stringsAsFactors`:
```{r}
grades <- data.frame(names = c("John", "Juan", "Jean", "Yao"), 
                     exam_1 = c(95, 80, 90, 85), 
                     exam_2 = c(90, 85, 85, 90),
                     stringsAsFactors = FALSE)
class(grades$names)
```
To convert a regular data frame to a tibble, you can use the `as_tibble` function.
```{r}
as_tibble(grades) %>% class()
```

## **5.11 The dot operator**
```{r}
tab_1 <- filter(murders, region == "South")
tab_2 <- mutate(tab_1, rate = total / population * 10^5)
rates <- tab_2$rate
median(rates)
```
`%>%`를 이용하면 object를 naming하는 것을 skip할 수 있다.
```{r}
filter(murders, region == "South") %>% 
  mutate(rate = total / population * 10^5) %>% 
  summarize(median = median(rate)) %>%
  pull(median)
```
For example to access the rate vector without the `pull` function we could use the dot operator:
```{r}
rates <-filter(murders, region == "South") %>% 
  mutate(rate = total / population * 10^5) %>% 
  .$rate
median(rates)
```

## **5.12 `do`**

The `do` functions serves as a bridge between R functions such as `quantile` and the tidyverse. The `do` function understands grouped tibbles and always returns a data frame.

We noted that if we attempt to use `quantile` to obtain the min, median and max in one call, we will receive an error: `Error: expecting result of length one, got : 2`.
```{r}
data(heights)
heights %>% 
  filter(sex == "Female") %>%
  summarize(range = quantile(height, c(0, 0.5, 1)))
```
We can use the `do` function fix this.
```{r}
my_summary <- function(dat){
  x <- quantile(dat$height, c(0, 0.5, 1))
  data_frame(min = x[1], median = x[2], max = x[3])
}
```
We can now apply the function to the heights dataset to obtain the summaries:
```{r}
heights %>% 
  group_by(sex) %>% 
  my_summary
```
We want a summary for each sex and the code returned just one summary. `do` makes this connection:
```{r}
heights %>% 
  group_by(sex) %>% 
  do(my_summary(.))
```
##### Note that here we need to use the dot operator. The tibble created by `group_by` is piped to `do`. Within the call to `do`, the name of this tibble is `.` and we want to send it to `my_summary`. If you do not use the dot, then `my_summary` has __no argument and returns an error telling us that `argument "dat"` is missing. You can see the error by typing:
```{r}
heights %>% 
  group_by(sex) %>% 
  do(my_summary())
```
This gives an error because do must always return a data frame. You can see the error by typing:
```{r}
heights %>% 
  group_by(sex) %>% 
  do(my_summary)
```

## **5.13 The purr package**
```{r}
compute_s_n <- function(n){
  x <- 1:n
  sum(x)
}
```
Using `sapply` to compute the sum of the first n integers for several values of n like this:
```{r}
n <- 1:25
s_n <- sapply(n, compute_s_n)
```
##### This type of operation, applying the same function or procedure to elements of an object, is quite common in data analysis. The **purrr** package includes functions similar to sapply but that better interact with other tidyverse functions. **purrr** functions will never do this: they will return objects of a specified type or return an error if this is not possible.

`map`, which works very similar to `sapply` but always, without exception, returns a list:
```{r}
library(purrr)
s_n <- map(n, compute_s_n)
class(s_n)
```
If we want a numeric vector, we can instead use `map_dbl` which always returns a vector of numeric values.
```{r}
s_n <- map_dbl(n, compute_s_n)
class(s_n)
```
A particularly useful **purrr** function for interacting with the rest of the tidyverse is `map_df`, which always returns a tibble data frame.
```{r}
compute_s_n <- function(n){
  x <- 1:n
  data_frame(sum = sum(x))
}
s_n <- map_df(n, compute_s_n)
head(s_n)
```

## ** 5.14 Tidyverse conditionals**

### **5.14.1 `case_when`**

The `case_when` function is useful for vectorizing conditional statements. It is similar to `ifelse` but can output any number of values, as opposed to just `TRUE` or `FALSE`.
```{r}
x <- c(-2, -1, 0, 1, 2)
case_when(x < 0 ~ "Negative", x > 0 ~ "Positive", TRUE ~ "Zero")
```
A common use for this function is to define categorical variables based on existing variables.
```{r}
data(murders)
murders %>% 
  mutate(group = case_when(
    abb %in% c("ME", "NH", "VT", "MA", "RI", "CT") ~ "New England",
    abb %in% c("WA", "OR", "CA") ~ "West Coast",
    region == "South" ~ "South",
    TRUE ~ "other")) %>%
  group_by(group) %>%
  summarize(rate = sum(total) / sum(population) * 10^5) %>%
  arrange(rate)
```

### **5.14.2 `between`**

A common operation in data analysis is to determine if a value falls inside an interval. We can check this using conditionals.

For example to check if the elements of a vector `x` are between `a` and `b` we can type:
```{r}
x >= a & x <= b
```
The `between` function performs the same operation.
```{r}
between(x, a, b)
```

## **5.15 Exercises**

**1. Load the `murders` dataset. Which of the following is true?**
```{r}
library(dslabs)
data(murders)
```
A. `murders` is in tidy format and is stored in a tibble.

**B. `murders` is in tidy format and is stored in a data frame.**

C. `murders` is not in tidy format and is stored in a tibble.

D. `murders` is not in tidy format and is stored in a data frame.

**2. Use `as_tibble` to covert the `murders` data table into a tibble and save it in an object called `murders_tibble`.
```{r}
murders_tibble <- as_tibble(murders)
```

**3. Use the `group_by` function to convert murders into a tibble that is grouped by region.**
```{r}
murders_tibble %>% group_by(region)
```

**4. Write tidyverse code that is equivalent to this code: `exp(mean(log(murders$population)))`. Write it using the pipe so that each function is called without arguments. Use the dot operator to access the population. Hint: The code should start with `murders %>%`.**
```{r}
murders %>%
  population %>%
  mean(log) %>%
  exp()
```

5. Use the `map_df` to create a data frame with three columns named `n`, `s_n`, and `s_n_2`. The first column should contain the numbers 1 through 100. The second and third columns should each contain the sum of 1 through *n* with *n* the row number.
```{r}
compute_s_n <- function(n){
  x <- 1:n
  data_frame(sum = sum(x))
}

s_n <- map_df(n, compute_s_n)
s_n_2 <- map_df(n, compute_s_n)

data.frame(n, s_n, s_n_2)
```

