---
title: "**Chapter 10 Data visualization in practice**"
output: html_notebook
---

## **10.1 Case study: new insights on poverty**

This section is based on two talks that exemplify this approach to education: New Insights on Poverty and The Best Stats You’ve Ever Seen. Specifically, in this section, we use data to attempt to answer the following two questions:

1. Is it a fair characterization of today’s world to say it is divided into western rich nations and the developing world in Africa, Asia and Latin America?

2. Has income inequality across countries worsened during the last 40 years?

To answer these questions, we will be using the `gapminder` dataset provided in **dslabs**.
```{r}
library(tidyverse)
library(dslabs)
data(gapminder)
gapminder %>% as_tibble()
```

### **10.1.1 Hans Rosling's quiz**

For each of the six pairs of countries below, which country do you think had the highest child mortality rates in 2015? Which pairs do you think are most similar?

1. Sri Lanka or Turkey

2. Poland or South Korea

3. Malaysia or Russia

4. Pakistan or Vietnam

5. Thailand or South Africa

When answering these questions without data, the non-European countries are typically picked as having higher child mortality rates: Sri Lanka over Turkey, South Korea over Poland, and Malaysia over Russia.

To answer these questions **with data**, we can use **dplyr**.
```{r}
gapminder %>% 
  filter(year == 2015 & country %in% c("Sri Lanka","Turkey")) %>% 
  select(country, infant_mortality)
```
We can use this code on all comparisons and find the following:
```{r}
gapminder %>% 
  filter(year == 2015 & country %in% c("Sri Lanka","Turkey","Poland","South Korea","Malaysia","Russia","Pakistan","Vietnam","Thailand","South Africa")) %>%
  select(country, infant_mortality) %>% arrange(infant_mortality)
```

## **10.2 Scatterplots**

In order to analyze this world view, our first plot is a scatterplot of life expectancy versus fertility rates (average number of children per woman). We start by looking at data from about 50 years ago, when perhaps this view was first cemented in our minds.
```{r}
filter(gapminder, year == 1962) %>%
  ggplot(aes(fertility, life_expectancy)) +
  geom_point()
```
Most points fall into two distinct categories:

1. Life expectancy around 70 years and 3 or less children per family.

2. Life expectancy lower then 65 years and more than 5 children per family.

To confirm that indeed these countries are from the regions we expect, we can use color to represent continent.
```{r}
filter(gapminder, year == 1962) %>%
  ggplot( aes(fertility, life_expectancy, color = continent)) +
  geom_point() 
```

## **10.3 Faceting**

We could easily plot the 2012 data in the same way we did for 1962. To make comparisons, however, side by side plots are preferable. In **ggplot2**, we can achieve this by *faceting* variables

To achieve faceting, we add a layer with the function `facet_grid`, which automatically separates the plots. This function lets you facet by up to two variables using columns to represent one variable and rows to represent the other. The function expects the row and column variables to be separated by a `~`. Here is an example of a scatterplot with `facet_grid` added as the last layer:
```{r}
filter(gapminder, year%in%c(1962, 2012)) %>%
  ggplot(aes(fertility, life_expectancy, col = continent)) +
  geom_point() +
  facet_grid(continent~year)
```
However, this is just an example and more than what we want, which is simply to compare 1962 and 2012. In this case, there is just one variable and we use `.` to let facet know that we are not using one of the variables:
```{r}
filter(gapminder, year%in%c(1962, 2012)) %>%
  ggplot(aes(fertility, life_expectancy, col = continent)) +
  geom_point() +
  facet_grid(.~year)
```

### **10.3.1 `facet_wrap`**

To explore how this transformation happened through the years, we can make the plot for several years. For example, we can add 1970, 1980, 1990, and 2000. If we do this, we will not want all the plots on the same row, the default behavior of `facet_grid`, since they will become too thin to show the data. Instead, we will want to use multiple rows and columns. The function `facet_wrap` permits us to do this by automatically wrapping the series of plots so that each display has viewable dimensions:
```{r}
years <- c(1962, 1980, 1990, 2000, 2012)
continents <- c("Europe", "Asia")
gapminder %>% 
  filter(year %in% years & continent %in% continents) %>%
  ggplot( aes(fertility, life_expectancy, col = continent)) +
  geom_point() +
  facet_wrap(~year) 
```

### **10.3.2 Fixed scales for better comparisons**

When using `facet`, this range is determined by the data shown in all plots and therefore kept fixed across plots. This makes comparisons across plots much easier. 

## **10.4 Time series plots**

Once we see these plots, new questions emerge. For example, which countries are improving more and which ones less? Was the improvement constant during the last 50 years or was it more accelerated during certain periods?

For a closer look that may help answer these questions, we introduce *time series* plots.
```{r}
gapminder %>% 
  filter(country == "United States") %>% 
  ggplot(aes(year, fertility)) +
  geom_point()
```
When the points are regularly and densely spaced, as they are here, we create curves by joining the points with lines, to convey that these data are from a single series, here a country. To do this, we use the `geom_line` function instead of `geom_point`.
```{r}
gapminder %>% 
  filter(country == "United States") %>% 
  ggplot(aes(year, fertility)) +
  geom_line()
```
This is particularly helpful when we look at two countries. If we subset the data to include two countries, one from Europe and one from Asia, then adapt the code above:
```{r}
countries <- c("South Korea","Germany")

gapminder %>% filter(country %in% countries) %>% 
  ggplot(aes(year,fertility)) +
  geom_line()
```
Rather than a line for each country, the points for both countries are joined. This is actually expected since we have not told `ggplot` anything about wanting two separate lines. To let `ggplot` know that there are two curves that need to be made separately, we assign each point to a `group`, one for each country:
```{r}
countries <- c("South Korea","Germany")

gapminder %>% filter(country %in% countries & !is.na(fertility)) %>%
  ggplot(aes(year, fertility, group = country)) +
  geom_line()
```
But which line goes with which country? We can assign colors to make this distinction. A useful side-effect of using the `color` argument to assign different colors to the different countries is that the data is automatically grouped:
```{r}
countries <- c("South Korea","Germany")

gapminder %>% filter(country %in% countries & !is.na(fertility)) %>%
  ggplot(aes(year,fertility, col = country)) +
  geom_line()
```

### **10.4.1 Labels instead of legends**

For trend plots **we recommend labeling the lines** rather than using legends since the viewer can quickly see which line is which country. 

We demonstrate how we can do this using the life expectancy data. We define a data table with the label locations and then use a second mapping just for these labels:
```{r}
labels <- data.frame(country = countries, x = c(1975,1965), y = c(60,72))

gapminder %>% 
  filter(country %in% countries) %>% 
  ggplot(aes(year, life_expectancy, col = country)) +
  geom_line() +
  geom_text(data = labels, aes(x, y, label = country), size = 5) +
  theme(legend.position = "none")
```

## **10.5 Data transformations**

First we learn how transformations can sometimes help provide more informative summaries and plots.

The `gapminder` data table includes a column with the countries gross domestic product (GDP). GDP measures the market value of goods and services produced by a country in a year. The GDP per person is often used as a rough summary of a country’s wealth. Here we divide this quantity by 365 to obtain the more interpretable measure *dollars* per day.
```{r}
gapminder <- gapminder %>%  mutate(dollars_per_day = gdp/population/365)
```

### **10.5.1 Log transformation**

Here is a histogram of per day incomes from 1970:
```{r}
past_year <- 1970
gapminder %>% 
  filter(year == past_year & !is.na(gdp)) %>%
  ggplot(aes(dollars_per_day)) + 
  geom_histogram(binwidth = 1, color = "black")
```
It might be more informative to quickly be able to see how many countries have average daily incomes of about $1 (extremely poor), $2 (very poor), $4 (poor), $8 (middle), $16 (well off), $32 (rich), $64 (very rich) per day. These changes are multiplicative and log transformations convert multiplicative changes into additive ones.
```{r}
gapminder %>% 
  filter(year == past_year & !is.na(gdp)) %>%
  ggplot(aes(log2(dollars_per_day))) + 
  geom_histogram(binwidth = 1, color = "black")
```

### **10.5.2 Which base?**

In general, we do not recommend using the natural log for data exploration and visualization.

For an example in which base 10 makes more sense, consider population sizes. A log base 10 is preferable since the range for these is:
```{r}
filter(gapminder, year == past_year) %>%
  summarize(min = min(population), max = max(population))
```
```{r}
gapminder %>% 
  filter(year == past_year) %>%
  ggplot(aes(log10(population))) +
  geom_histogram(binwidth = 0.5, color = "black")
```

### **10.5.3 Transform the values or the scale?

There are two ways we can use log transformations in plots. We can log the values before plotting them or use log scales in the axes. Both approaches are useful and have different strengths. If we log the data, we can more easily interpret intermediate values in the scale.

As we learned earlier, if we want to scale the axis with logs, we can use the `scale_x_continuous` function. Instead of logging the values first, we apply this layer:
```{r}
gapminder %>% 
  filter(year == past_year & !is.na(gdp)) %>%
  ggplot(aes(dollars_per_day)) + 
  geom_histogram(binwidth = 1, color = "black") +
  scale_x_continuous(trans = "log2")
```
이 경우에는 지수로 표현되지 않고 지수를 포함한 whole number가 나온다.

## **10.6 Visualizing multimodal distributions

In the histogram above we see two *bumps*: one at about 4 and another at about 32. In statistics these bumps are sometimes referred to as *modes*. The mode of a distribution is the value with the highest frequency. The mode of the normal distribution is the average.

## **10.7 Comparing multiple distributions with boxplots and ridge plots**

However, the histogram does not show us if the two groups of countries are *west* versus the *developing* world.
```{r}
p <- gapminder %>% 
  filter(year == past_year & !is.na(gdp)) %>%
  ggplot(aes(region, dollars_per_day)) 
p + geom_point() 
```
Notice that we can’t read the region names because the default **ggplot2** behavior is to write the labels horizontally and here we run out of room. We can easily fix this by either rotating the axes or rotating the labels. Consulting the cheat sheet we find we can rotate the axes with `coord_flip` or we can rotate the labels by changing the `theme` through `element_text`. The `hjust = 1` justifies the text so that it is next to the axis.
```{r}
p + geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
In general, **do not order alphabetically**. Data exploration is often made easier if we order the strata by some interpretable quantity. This can be achieved using the `reorder` function.
```{r}
p <- gapminder %>% 
  filter(year == past_year & !is.na(gdp)) %>%
  mutate(region = reorder(region, dollars_per_day, FUN = median)) %>%
  ggplot(aes(region, dollars_per_day)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
p
```
As with the histogram, if we remake the plot using a log scale. This permits us to better appreciate the differences within the developing world.
```{r}
p + scale_y_continuous(trans = "log2")
```

### **10.7.1 Boxplots**

Using a histogram, we found a bimodal distribution with the modes relating to poor and rich countries.
```{r}
gapminder <- gapminder %>% 
  mutate(group = case_when(
    region %in% c("Western Europe", "Northern Europe","Southern Europe", 
                    "Northern America", "Australia and New Zealand") ~ "West",
    region %in% c("Eastern Asia", "South-Eastern Asia") ~ "East Asia",
    region %in% c("Caribbean", "Central America", "South America") ~ "Latin America",
    continent == "Africa" & region != "Northern Africa" ~ "Sub-Saharan Africa",
    TRUE ~ "Others"))
```
We turn this `group` variable into a factor to control the order of the levels:
```{r}
gapminder <- gapminder %>% 
  mutate(group = factor(group, 
                        levels = c("Others", "Latin America", "East Asia", "Sub-Saharan Africa", "West")))
```
We now want to compare the distribution across these five groups to confirm the “west versus the rest” dichotomy.
```{r}
p <- gapminder %>% 
  filter(year == past_year & !is.na(gdp)) %>%
  ggplot(aes(group, dollars_per_day)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_continuous(trans = "log2") +
  xlab("")
p 
```
Boxplots have the limitation that by summarizing the data into five numbers, we might miss important characteristics of the data. One way to avoid this is by showing the data.
```{r}
p + geom_point(alpha = 0.5)
```

### **10.7.2 Ridge plots**

In cases in which we are concerned that the boxplot summary is too simplistic, we can show stacked smooth densities or histograms. We refer to these as *ridge plots*. Because we are used to visualizing densities with values in the x-axis, we stack them vertically.
```{r}
library(ggridges)
p <- gapminder %>% 
  filter(year == past_year & !is.na(dollars_per_day)) %>%
  ggplot(aes(dollars_per_day, group)) + 
  scale_x_continuous(trans = "log2") 
p  + geom_density_ridges() 
```
Note that we have to invert the `x` and `y` used for the boxplot. A useful `geom_density_ridges` parameter is `scale`, which lets you determine the amount of overlap, with `scale = 1` meaning no overlap and larger values resulting in more overlap.
```{r}
p + geom_density_ridges(jittered_points = TRUE)
```
To show data points, but without using jitter we can use the following code to add what is referred to as a *rug representation* of the data.
```{r}
p + geom_density_ridges(jittered_points = TRUE, 
                        position = position_points_jitter(width = 0.05, height = 0),
                        point_shape = '|', point_size = 3, point_alpha = 1, alpha = 0.7)
```

### **10.7.3 Example: 1970 versus 2010 income distributions**

```{r}
past_year <- 1970
present_year <- 2010
gapminder %>% 
  filter(year %in% c(past_year, present_year) & !is.na(gdp)) %>%
  mutate(west = ifelse(group == "West", "West", "Developing")) %>%
  ggplot(aes(dollars_per_day)) +
  geom_histogram(binwidth = 1, color = "black") +
  scale_x_continuous(trans = "log2") + 
  facet_grid(year ~ west)
```
We remake the plots using only countries with data available for both years. In the data wrangling part of this book, we will learn **tidyverse** tools that permit us to write efficient code for this, but here we can use simple code using the `intersect` function:
```{r}
country_list_1 <- gapminder %>% 
  filter(year == past_year & !is.na(dollars_per_day)) %>% 
  pull(country)

country_list_2 <- gapminder %>% 
  filter(year == present_year & !is.na(dollars_per_day)) %>% 
  pull(country)
      
country_list <- intersect(country_list_1, country_list_2)
```
Let’s remake the plot, but only for this subset by simply adding `country %in% country_list` to the filter function:
```{r}
past_year <- 1970
present_year <- 2010
gapminder %>% 
  filter(year %in% c(past_year, present_year) & !is.na(gdp) & country %in% country_list) %>%
  mutate(west = ifelse(group == "West", "West", "Developing")) %>%
  ggplot(aes(dollars_per_day)) +
  geom_histogram(binwidth = 1, color = "black") +
  scale_x_continuous(trans = "log2") + 
  facet_grid(year ~ west)
```
To see which specific regions improved the most, we can remake the boxplots we made above, but now adding the year 2010 and then using facet to compare the two years.
```{r}
gapminder %>% 
  filter(year %in% c(past_year, present_year) & country %in% country_list) %>%
  ggplot(aes(group, dollars_per_day)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_continuous(trans = "log2") +
  xlab("") +
  facet_grid(. ~ year)
```
So instead of faceting, we keep the data from each year together and ask to color (or fill) them depending on the year. Note that groups are automatically separated by year and each pair of boxplots drawn next to each other.
```{r}
gapminder %>% 
  filter(year %in% c(past_year, present_year) & country %in% country_list) %>%
  mutate(year = factor(year)) %>%
  ggplot(aes(group, dollars_per_day, fill = year)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_continuous(trans = "log2") +
  xlab("") 
```
Let’s start by noting that density plots for income distribution in 1970 and 2010 deliver the message that the gap is closing:
```{r}
gapminder %>% 
  filter(year %in% c(past_year, present_year) & country %in% country_list) %>%
  ggplot(aes(dollars_per_day)) +
  geom_density(fill = "grey") + 
  scale_x_continuous(trans = "log2") + 
  facet_grid(. ~ year)
```
But when we overlay two densities, the default is to have the area represented by each distribution add up to 1, regardless of the size of each group:
```{r}
gapminder %>% 
  filter(year %in% c(past_year, present_year) & country %in% country_list) %>%
  mutate(group = ifelse(group == "West", "West", "Developing")) %>%
  ggplot(aes(dollars_per_day, fill = group)) +
  scale_x_continuous(trans = "log2") +
  geom_density(alpha = 0.2) + 
  facet_grid(year ~ .)
```

### **10.7.4 Accessing computed variables**

In **ggplot2**, we access these variables by surrounding the name with two dots `..`. We will use the following mapping:
```{r}
aes(x = dollars_per_day, y = ..count..)
```

```{r}
p <- gapminder %>% 
  filter(year %in% c(past_year, present_year) & country %in% country_list) %>%
  mutate(group = ifelse(group == "West", "West", "Developing")) %>%
  ggplot(aes(dollars_per_day, y = ..count.., fill = group)) +
  scale_x_continuous(trans = "log2", limit = c(0.125, 300))

p + geom_density(alpha = 0.2) + 
  facet_grid(year ~ .)
```
If we want the densities to be smoother, we use the `bw` argument so that the same bandwidth is used in each density. We selected 0.75 after trying out several values.
```{r}
p + geom_density(alpha = 0.2, bw = 0.75) + facet_grid(year ~ .)
```
To visualize if any of the groups defined above are driving this we can quickly make a ridge plot:
```{r}
gapminder %>% 
  filter(year %in% c(past_year, present_year) & !is.na(dollars_per_day)) %>%
  ggplot(aes(dollars_per_day, group)) + 
  scale_x_continuous(trans = "log2") + 
  geom_density_ridges(adjust = 1.5) +
  facet_grid(. ~ year)
```
Another way to achieve this is by stacking the densities on top of each other:
```{r}
gapminder %>% 
    filter(year %in% c(past_year, present_year) & country %in% country_list) %>%
  group_by(year) %>%
  mutate(weight = population/sum(population)*2) %>%
  ungroup() %>%
  ggplot(aes(dollars_per_day, fill = group)) +
  scale_x_continuous(trans = "log2", limit = c(0.125, 300)) + 
  geom_density(alpha = 0.2, bw = 0.75, position = "stack") + facet_grid(year ~ .)
```

### **10.7.5 Weighted densities**

We can actually weight the smooth densities using the `weight` mapping argument.


### **10.8 The ecological fallacy and importance of showing the data**

We start by comparing these quantities across regions, but before doing this, we define a few more regions:
```{r}
gapminder <- gapminder %>% 
  mutate(group = case_when(
    region %in% c("Western Europe", "Northern Europe","Southern Europe", 
                    "Northern America", "Australia and New Zealand") ~ "West",
    region %in% "Northern Africa" ~ "Northern Africa",
    region %in% c("Eastern Asia", "South-Eastern Asia") ~ "East Asia",
    region == "Southern Asia"~ "Southern Asia",
    region %in% c("Central America", "South America", "Caribbean") ~ "Latin America",
    continent == "Africa" & region != "Northern Africa" ~ "Sub-Saharan Africa",
    region %in% c("Melanesia", "Micronesia", "Polynesia") ~ "Pacific Islands"))
```
We then compute these quantities for each region:
```{r}
surv_income <- gapminder %>% 
  filter(year %in% present_year & !is.na(gdp) & 
           !is.na(infant_mortality) & !is.na(group)) %>%
  group_by(group) %>%
  summarize(income = sum(gdp)/sum(population)/365,
            infant_survival_rate = 
              1 - sum(infant_mortality/1000*population)/sum(population)) 

surv_income %>% arrange(income)
```
This shows a dramatic difference. While in the West less than 0.5% of infants die, in Sub-Saharan Africa the rate is higher than 6%! The relationship between these two variables is almost perfectly linear:
```{r}
surv_income %>% 
  ggplot(aes(income, infant_survival_rate, label=group, color=group)) +
  scale_x_continuous(trans = "log2") +
  geom_label(size=3)
```

### **10.8.1 Logistic transformation**

The logistic or logit transformation for a proportion or rate *p* is defined as:
$$ f(p)=log(p/1-p) $$
### **10.8.2 Show the data**

Jumping to this conclusion based on a plot showing averages is referred to as the *ecological fallacy*. The almost perfect relationship between survival rates and income is only observed for the averages at the region level.
```{r}
gapminder <- gapminder %>% 
  mutate(infant_survival_rate = 
              1 - infant_mortality/1000)

gapminder %>%
  filter(year == past_year & !is.na(group)) %>%
  ggplot(aes(dollars_per_day, infant_survival_rate, color=group)) +
  scale_x_continuous(trans = "log2", limit = c(0.25, 100)) +
  geom_point(alpha=0.5, size=2)
```

