---
title: "**Chapter 12 Robust summaries**"
output: html_notebook
---

## **12.1 Outliers**

Here we discuss outliers, approaches that can help detect them, and summaries that take into account their presence. How do we distinguish an outlier from measurements that were too big or too small simply due to expected variability? This is not always an easy question to answer, but we try to provide some guidance.

```{r}
library(dslabs)
data(outlier_example)
str(outlier_example)
```

```{r}
mean(outlier_example)
sd(outlier_example)
```

```{r}
qplot(outlier_example)
```

There appears to be at least one value that is nonsensical, since we know that a height of 180 feet is impossible. The boxplot detects this point as an outlier:

```{r}
boxplot(outlier_example)
```

## **12.2 Median**

The median, defined as the value for which half the values are smaller and the other half are bigger, is robust to such outliers. No matter how large we make the largest point, the median remains the same.

```{r}
median(outlier_example)
```

## **12.3 The inter quartile range(IQR)**

The box in boxplots are defined by the first and third quartile. These are meant to provide an idea of the variability in the data: 50% of the data is within this range.

We can do some math to see that for normally distributed data, the IQR / 1.349 approximates the standard deviation of the data had an outlier not been present.
 
```{r}
IQR(outlier_example) / 1.349 
```

## **12.4 Tukey's definition of an outlier**

In R, points falling outside the whiskers of the boxplot are referred to as outliers. The top whisker ends at the 75th percentile plus 1.5×IQR. Similarly the bottom whisker ends at the 25th percentile minus 1.5×IQR.

$$ [Q1-1.5×(Q3-Q1), Q3+1.5×(Q3-Q1)] $$

```{r}
q3 <- qnorm(0.75)
q1 <- qnorm(0.25)
iqr <- q3 - q1
r <- c(q1 - 1.5*iqr, q3 + 1.5*iqr)
r
```

Using the `pnorm` function, we see that 99.3% of the data falls in this interval. If we want an outlier to be rarer, we can increase the 1.5 to a larger number. Tukey also used 3 and called these far out outliers. With a normal distribution, 100% of the data falls in this interval. This translates into about 2 in a million chance of being outside the range. In the `geom_boxplot` function, this can be controlled by the `outlier.size` argument, which defaults to 1.5.

```{r}
max_height <- quantile(outlier_example, 0.75) + 3*IQR(outlier_example)
max_height
```

If we take this value out, we can see that the data is in fact normally distributed as expected:

```{r}
x <- outlier_example[outlier_example < max_height]
qqnorm(x)
qqline(x)
```

## **12.5 Median absolute deviation**

Another way to robustly estimate the standard deviation in the presence of outliers is to use the median absolute deviation (MAD). To compute the MAD, we first compute the median, and then for each value we compute the distance between that value and the median. The MAD is defined as the median of these distances.

```{r}
mad(outlier_example)
```

## **12.6 Exercises**

```{r}
install.packages("HistData")
library(HistData)
data(Galton)
x <- Galton$child
```

**1. Compute the average and median of these data.**
```{r}
mean(x)
median(x)
```

**2. Compute the median and median absolute deviation of these data.**
```{r}
median(x)
mad(x)
```

**3. Now suppose Galton made a mistake when entering the first value and forgot to use the decimal point. You can imitate this error by typing:**
```{r}
x_with_error <- x
x_with_error[1] <- x_with_error[1]*10
```

**How many inches does the average grow after this mistake?**
```{r}
mean(x)
mean(x_with_error)
```

**4. How many inches does the SD grow after this mistake?**
```{r}
sd(x)
sd(x_with_error)
```

**5. How many inches does the median grow after this mistake?**
```{r}
median(x)
median(x_with_error)
```

**6. How many inches does the MAD grow after this mistake?**
```{r}
mad(x)
mad(x_with_error)
```

**7. How could you use exploratory data analysis to detect that an error was made?**

A. Since it is only one value out of many, we will not be able to detect this.

B. We would see an obvious shift in the distribution.

**C. A boxplot, histogram, or qq-plot would reveal a clear outlier.**

D. A scatter plot would show high levels of measurement error.

**8. How much can the average accidentally grow with mistakes like this? Write a function called `error_avg` that takes a value `k` and returns the average of the vector `x` after the first entry changed to `k`. Show the results for `k=10000` and `k=-10000`.**
```{r}
error <- x
error[1] <- 10000
error_avg <- mean(error)
error_avg

error <- x
error[1] <- -10000
error_avg <- mean(error)
error_avg
```

## **12.7 Case study: self-reported student heights**

```{r}
library(dslabs)
data("reported_heights")
head(reported_heights)
```
Height is a character vector so we create a new column with the numeric version:
```{r}
reported_heights <- reported_heights %>%
  mutate(original_heights = height, height = as.numeric(height))
```

Note that we get a warning about NAs. This is because some of the self reported heights were not numbers.
```{r}
reported_heights %>% 
  filter(is.na(height)) %>% 
  head()
```

Some students self-reported their heights using feet and inches rather than just inches. Others used centimeters and others were just trolling. For now we will remove these entries:
```{r}
reported_heights <- filter(reported_heights, !is.na(height))
```

```{r}
reported_heights %>% 
  group_by(sex) %>%
  summarize(average = mean(height), sd = sd(height),
            median = median(height), MAD = mad(height))
```

This suggests that we have outliers, which is confirmed by creating a boxplot:
```{r}
reported_heights %>% group_by(sex) %>% 
  ggplot(aes(sex, height)) + geom_boxplot()
```

We can see some rather extreme values. To see what these values are, we can quickly look at the largest values using the `arrange` function:
```{r}
reported_heights %>% 
  arrange(desc(height)) %>% 
  top_n(10, height)
```

We can review all the nonsensical answers by looking at the data considered to be far out by Tukey:
```{r}
max_height <- quantile(reported_heights$height, .75) + 3*IQR(reported_heights$height)
min_height <- quantile(reported_heights$height, .25) - 3*IQR(reported_heights$height)
c(min_height, max_height)
```

```{r}
reported_heights %>% 
  filter(!between(height, min_height, max_height)) %>% 
  select(original_heights) %>%
  head(n=10)
```
Examining these heights carefully, we see two common mistakes: entries in centimeters, which turn out to be too large, and entries of the form `x.y` with `x` and `y` representing feet and inches respectively, which turn out to be too small. Some of the even smaller values, such as 1.6, could be entries in meters.

In the Data Wrangling part of this book we will learn techniques for correcting these values and converting them into inches.
