---
title: "Chapter 4"
output: html_notebook
---

## 4.4 For-loops

1+2+...+n=n(n+1)/2 이라는 식이 확실하지 않다면 어떻게 확인을 할까?
```{r}
compute_s_n <- function(n){
  x <- 1:n
  sum(x)
}
```

*Sn*을 계산할 때 만약 n값을 1부터 25까지 지정한다면, 25줄을 다 써야할까? `For-loops` let us define the range that our variable takes, then change the value and evaluate expression as you loop.
```{r}
for(i in 1:5){
  print(i)
}
```

Here is the for-loop we would write for our *Sn* example:
```{r}
m <- 25
s_n <- vector(length=m) # create an empty vector
for(n in 1:m){
  s_n[n] <- compute_s_n(n)
}
```

이를 plot을 이용해 확인할 수 있다.
```{r}
n <- 1:m
plot(n, s_n)
```

```{r}
head(data.frame(s_n = s_n, formula = n*(n+1)/2))
```

두 결과를 겹쳐서 보고 싶으면 `lines`를 이용해서 그릴 수 있다.
```{r}
plot(n, s_n)
lines(n, n*(n+1)/2)
```

## 4.5 Vectorization and functionals

```{r}
x <- 1:10
sqrt(x)

y <- 1:10
x*y
```

`Seqeunce`로 지정을 했을 때 각 수를 하나의 element로 보는 것이 아닌 sequence를 scalar로 보기 때문에 sequence의 첫번재 수만 인식하여 `compute_s_n`에 사용한다.
```{r}
n <- 1:25
compute_s_n(n)
```

The function `sapply` permits us to perform element-wise operations on any function.
```{r}
x <- 1:10
sapply(x, sqrt)
```
결과적으로 `sapply`를 이용하면 지정한 값을 element-wise로 인식한다. 따라서 앞서 사용했던 compute_s_n에 `sapply`를 이용하면 기대하던 값을 얻을 수 있다.
```{r}
n <- 1:25
s_n <- sapply(n, compute_s_n)
plot(n, s_n)
```

## 4.6 Exercises

1. What will this conditional expression return?
```{r}
x <- c(1,2,-3,4)

if(all(x>0)){
  print("All Postives")
} else{
  print("Not all positives")
}
```

2. Which of the following expressions is always `FALSE` when at least one entry of a logical vector `x` is TRUE?
```{r}
x <- c(TRUE, TRUE, FALSE)
all(x)
any(x)
any(!x)
all(!x)
```
A. `all(x)`
B. `any(x)`
C. `any(!x)`
**D. `all(!x)`**

3. The function `nchar` tells you how many characters long a character vector is.
Write a line of code that assigns to the object `new_names` the state abbreviation when the state name is longer than 8 characters.
```{r}
state_8 <- nchar(murders$state) > 8
new_names <- murders$abb[state_8]
new_names
```

4. Create a function `sum_n` that for any given value, say *n*, computes the sum of the integers from 1 to n (inclusive). Use the function to determine the sum of integers from 1 to 5,000.
```{r}
sum_n <- function(n){
  x <- 1:n
  sum(x)
}
sum_n(5000)
```

5. Create a function `altman_plot` that takes two arguments, `x` and `y`, and plots the difference against the sum.
```{r}

```

6. After running the code below, what is the value of `x`? 
```{r}
x <- 3
my_func <- function(y){
  x <- 5
  y+5
}

print(x)
```
x의 value는 3인 것을 알 수 있다. 

7. Write a function `compute_s_n` that for any given n computes the sum Sn=1^2+2^2+3^2+...+n^2. Report the value of the sum when n=10
```{r}
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}
compute_s_n(10)
```

8. Define an empty numerical vector `s_n` of size 25 using `s_n <- vector("numeric", 25)` and store in the results of S1, S2, ... S25 using a for-loop.
```{r}
m <- 25
s_n <- vector("numeric", 25)
for(n in 1:m){
  s_n[n] <- compute_s_n(n)
}

s_n
```


9. Repeat exercise 8, but this time use `sapply`.
```{r}
n <- 1:25
s_n <- sapply(n, compute_s_n)

s_n
```

10. Repeat exercise 8, but this time use `map_dbl`
```{r}
n <- 1:25
s_n <- map_dbl(n, compute_s_n)

s_n
```

11. Plot *Sn* versus *n*. Use points defined by n = 1,...,25.
```{r}
n <- 1:25
plot(n, s_n)
```

12. Confirm that the formula for this sum is Sn=n(n+1)(2n+1)/6
```{r}
s_n = s_n
formula = n*(n+1)*(2n+1)/6
s_n == formula
```
?error 뜨는 이유
