---
title: "Chapter 5 The tidyverse"
output: html_notebook
---

Chapter 5는 `tidyverse` package를 이용해 진행된다.
```{r}
install.packages('tidyverse')
library(tidyverse)
```

## 5.1 Tidy data
Row: one observation

Column: the different variables available for each of these observations

`murders` dataset과 같은 것을 **tidy** data frame이라고 한다.
```{r}
#>        state abb region population total
#> 1    Alabama  AL  South    4779736   135
#> 2     Alaska  AK   West     710231    19
#> 3    Arizona  AZ   West    6392017   232
#> 4   Arkansas  AR  South    2915918    93
#> 5 California  CA   West   37253956  1257
#> 6   Colorado  CO   West    5029196    65
```

## 5.2 Exercises

1. Examine the built-in dataset `co2`. Which of the following is true:
```{r}
data("co2")
co2
```
A. `co2` is tidy data: it has one year for each row.

B. `co2` is not tidy: we need at least one column with a character vector.

**C. `co2` is not tidy: it is a matrix instead of a data frame.**

D. `co2` is not tidy: to be tidy we would have to wrangle it to have three columns (year, month and value), then each co2 observation would have a row.

2. Examine the built-in dataset `ChickWeight`. Which of the following is true:
```{r}
data("ChickWeight")
ChickWeight
```
A. `ChickWeight` is not tidy: each chick has more than one row.

**B. `ChickWeight` is tidy: each observation (a weight) is represented by one row. The chick from which this measurement came from is one the variables.**

C. `ChickWeight` is not a tidy: we are missing the year column.

D. `ChickWeight` is tidy: it is stored in a data frame.

3. Examine the built-in dataset `BOD`. Which of the following is true:
```{r}
data("BOD")
BOD
```
A. BOD is not tidy: it only has six rows.

B. BOD is not tidy: the first column is just an index.

**C. BOD is tidy: each row is an observation with two values (time and demand)**

D. BOD is tidy: all small datasets are tidy by definition.

4. Which of the following built-in datasets is tidy (you can pick more than one):
```{r}
data("BJsales")
data("EuStockMarkets")
data("DNase")
data("Formaldehyde")
data("Orange")
data("UCBAdmissions")
```
A. `BJsales`

**B. `EuStockMarkets`**

**C. `DNase`**

**D. `Formaldehyde`**

**E. `Orange`**

F. `UCBAdmissions`

## 5.3 Manipulating data frames

To change the data table by adding a new column, we use `mutate`. To filter the data table to a subset of rows, we use `filter`. Finally, to subset the data by selecting specific columns, we use `select`.

### 5.3.1 Adding a column with `mutate`

The function `mutate` takes the data frame as a first argument and the name and values of the variable as a second argument using the convention `name = values`. 
```{r}
library(dslabs)
data("murders")
murders <- mutate(murders, rate = total / population * 100000)
```
`total`과 `population`이 따로 define되지 않았으나 error가 뜨지 않는 이유: Functions in this package, such as mutate, know to look for variables in the data frame provided in the first argument.

```{r}
head(murders)
```

### 5.3.2 Subsetting with `filter`

Now suppose that we want to filter the data table to only show the entries for which the murder rate is lower than 0.71. To do this we use the `filter` function, which takes the data table as the first argument and then the conditional statement as the second.
```{r}
filter(murders, rate <= 0.71)
```

### 5.3.3 Selecting columns with `select`

Although our data table only has six columns, some data tables include hundreds. If we want to view just a few, we can use the **dplyr** `select` function.
```{r}
new_table <- select(murders, state, region, rate)
filter(new_table, rate <= 0.71)
```

## 5.4 Exercises

1. Load the **dplyr** package and the murders dataset.
```{r}
library(dplyr)
library(dslabs)
data(murders)
```
You can add columns using the **dplyr** function `mutate`. 
```{r}
murders <- mutate(murders, population_in_millions = population / 10^6)
```
Use the function `mutate` to add a murders column named `rate` with the per 100,000 murder rate as in the example code above. Make sure you redefine `murders` as done in the example code above ( murders <- [your code]) so we can keep using this variable.
```{r}
murders <- mutate(murders, rate = total / population * 10^5)
head(murders)
```

2. If `rank(x)` gives you the ranks of `x` from lowest to highest, `rank(-x)` gives you the ranks from highest to lowest. Use the function `mutate` to add a column `rank` containing the rank, from highest to lowest murder rate. Make sure you redefine `murders` so we can keep using this variable.
```{r}
murders <- mutate(murders, rank = rank(-rate))
head(murders)
```

3. With **dplyr**, we can use `select` to show only certain columns. For example, with this code we would only show the states and population sizes:
```{r}
select(murders, state, population) %>% head()
```
Use `select` to show the state names and abbreviations in `murders`. Do not redefine `murders`, just show the results.
```{r}
select(murders, state, abb) %>% head()
```

4. The **dplyr** function `filter` is used to choose specific rows of the data frame to keep. Unlike `select` which is for columns, `filter` is for rows. For example, you can show just the New York row like this:
```{r}
filter(murders, state == "New York")
```
Use `filter` to show the top 5 states with the highest murder rates. After we add murder rate and rank, do not change the murders dataset, just show the result. Remember that you can filter based on the `rank` column.
```{r}
filter(murders, rank <= 5)
```

5. We can remove rows using the `!=` operator. For example, to remove Florida, we would do this:
```{r}
no_florida <- filter(murders, state != "Florida")
```
Create a new data frame called `no_south` that removes states from the South region. How many states are in this category? You can use the function `nrow` for this.
```{r}
no_south <- filter(murders, region != "South")
nrow(no_south)
```

6. We can also use `%in%` to filter with **dplyr**. You can therefore see the data from New York and Texas like this:
```{r}
filter(murders, state %in% c("New York", "Texas"))
```
Create a new data frame called murders_nw with only the states from the Northeast and the West. How many states are in this category?
```{r}
murders_nw <- filter(murders, region %in% c("Northeast", "West"))
nrow(murders_nw)
```

7. Suppose you want to live in the Northeast or West **and** want the murder rate to be less than 1. We want to see the data for the states satisfying these options. Note that you can use logical operators with `filter`. Here is an example in which we filter to keep only small states in the Northeast region.
```{r}
filter(murders, population < 5000000 & region == "Northeast")
```
Make sure `murders` has been defined with `rate` and `rank` and still has all states. Create a table called `my_states` that contains rows for states satisfying both the conditions: it is in the Northeast or West and the murder rate is less than 1. Use `select` to show only the state name, the rate and the rank.
```{r}
my_states <- filter(murders, region %in% c("Northeast", "West") & rate < 1)
select(my_states, state, rate, rank)
```

## 5.5 The pipe: `%>%`

We can write code that looks more like a description of what we want to do without intermediate objects:

*original data* -> *select* -> *filter*

For such an operation, we can use the pipe `%>%`.
```{r}
murders %>% select(state, region, rate) %>% filter(rate <= 0.71)
```
In general, the pipe sends the result of the left side of the pipe to be the first argument of the function on the right side of the pipe.
```{r}
#For example,
16 %>% sqrt()
16 %>% sqrt() %>% log(base=2)
```

## 5.6 Exercises

1. The pipe `%>%` can be used to perform operations sequentially without having to define intermediate objects. Start by redefining murder to include rate and rank.
```{r}
murders <- mutate(murders, rate =  total / population * 100000, rank = rank(-rate))
```
In the solution to the previous exercise, we did the following:
```{r}
my_states <- filter(murders, region %in% c("Northeast", "West") & rate < 1)
select(my_states, state, rate, rank)
```
The pipe `%>%` permits us to perform both operations sequentially without having to define an intermediate variable `my_states`. We therefore could have mutated and selected in the same line like this:
```{r}
mutate(murders, rate =  total / population * 100000, rank = rank(-rate)) %>% select(state, rate, rank)
```
Notice that `select` no longer has a data frame as the first argument. The first argument is assumed to be the result of the operation conducted right before the `%>%`.

Repeat the previous exercise, but now instead of creating a new object, show the result and only include the state, rate, and rank columns. Use a pipe `%>%` to do this in just one line.
```{r}
murders %>% filter(region %in% c("Northeast", "West") & rate < 1) %>% select(state, rate, rank)
```

2. Reset `murders` to the original table by using `data(murders)`. Use a pipe to create a new data frame called `my_states` that considers only states in the Northeast or West which have a murder rate lower than 1, and contains only the state, rate and rank columns. The pipe should also have four components separated by three `%>%`. The code should look something like this:
```{r}
my_states <- murders %>%
  mutate SOMETHING %>% 
  filter SOMETHING %>% 
  select SOMETHING
```

```{r}
data("murders")
my_states <- murders %>%
  mutate(rate =  total / population * 10^5, rank = rank(-rate)) %>%
  filter(region %in% c("Northeast", "West") & rate < 1) %>%
  select(state, rate, rank)
my_states
```

